---
title: "Exploring the Influence of Lifestyle Factors on Obesity Prevelance in Scotland"
format: html
editor: visual
execute:
  echo: false
self-contained: true
toc: true
number-sections: true
editor_options: 
  chunk_output_type: inline
---

```{r}
#| echo: false
#| warning: false
#| message: false 
library(sjPlot)
library(ggplot2)
library(dplyr)
library(moderndive)
library(skimr)
library(kableExtra)
library(gridExtra)
library(MASS)
library(readr)
library(tidyr)
library(gt)
library(tidyverse)
library(tidymodels)
library(patchwork)
```

# Introduction

Time and time again in the UK we see obesity putting more people's lives at risk and more pressure on the NHS. Not only does obesity cause majority health complications but it also puts strain on the NHS's already dwindling budget. So the question arises, is our government doing enough to combat this epidemic?

There is no one solution to cure obesity as many different socio-economic factors play a role in a person's diet and lifestyle. In the following report we shall investigate the change in the prevalence of obesity in Scotland from 2008-2012, as well as determining if obesity rates are influenced by ages, gender, socio-economic status or other lifestyle factor.

We shall conduct a brief exploratory analysis in @sec-EA followed by more rigorous analysis in @sec-FA and with concluding remarks in @sec-C.

# Exploratory Analysis {#sec-EA}

```{r}
#| echo: false
#| warning: false
#Creating the binary variable for obesity

data <- read.csv('DAProject9.csv')
library(dplyr)
data <- data %>%
  mutate(Obese = ifelse(BMIgroup == "Obese", 1, 0))

```

First, we created a new indicator covariate "Obese" according to the original variable "BMIgroup". As we can see, it was a categorical response variable with four different levels. Our aim was to explore the impact of different factors on obesity, so we use 1 and 0 to separate obese people and other type of people.

```{r}
#| echo: false
#| warning: false
#Attempting to create the time series plot
data_2008 <- subset(data, Year == '2008' & Obese == 1)
data_2009 <- subset(data, Year == '2009'& Obese == 1)
data_2010 <- subset(data, Year == '2010'& Obese == 1)
data_2011 <- subset(data, Year == '2011'& Obese == 1)
data_2012 <- subset(data, Year == '2012'& Obese == 1)


year <- c(2008, 2009, 2010, 2011, 2012)
no_obese <- c(1448, 1647, 1666, 1620, 1088)

year_obese <- cbind(year, no_obese)
  year_obese <- as.data.frame(year_obese)

```

Second, we calculated the value of obesity number by year and found each year the population sizes were different. In order to avoid the influence of population size, we calculated the proportion of prevalence of obesity over year and drew a time series plot of obese observations from 2008 to 2012 to show whether the proportion of prevalence of obesity in Scotland changed over the years.

```{r}
#| echo: false
#| warning: false
#| fig-cap: Time Series Plot of Obese Observations Over Time.
#| label: fig-time
#| fig-align: center
#| fig-width: 4.5
#| fig-height: 3.5

library(ggplot2)

#time series plot with proportion of obesity
obesity_by_time=data|>group_by(Year)|>
  summarise(Obesity_Prevalence=mean(Obese==1)*100)
ggplot(obesity_by_time,aes(x=Year,y=Obesity_Prevalence))+
  geom_line()+
  labs(x="Year",y="Percentage of the density")
```

From @fig-time , we could see that the obesity prevalence proportion first decreased from 2008 to 2009 and got the lowest value 28.78%. Then it increased dramatically and reached its peak value, 30.47% , in 2010. After 2010, it kept decreasing and it got 29.61% in 2012.

```{r}
#| echo: false
#| warning: false
#| tbl-cap: Obese Counts Per Year
#| label: tbl-counts

#Summary table
summary_by_year <- data %>%
  group_by(Year) %>%
  summarize(Count_Obese = sum(Obese))
gt(summary_by_year)
```

From @tbl-counts we see the overall counts of obese observations per year. Again this is helpful to visualise the obese observations however, as stated above, proportions were preferred for this study as it was not balanced.

```{r}
#| echo: false
#| warning: false
#| fig-cap: Proportion of obesity by age groups.
#| label: fig-scat1
#| fig-align: center
#| fig-width: 4.5
#| fig-height: 3.5


#barplots of obesity on each covariate
library(dplyr)
# age
obese_age.freq=table(data$AgeGroup)
obese_age.freq=as.data.frame(obese_age.freq)
colnames(obese_age.freq)=c("obese_age","number")
obese_age.freq$proportion=obese_age.freq$number/sum(obese_age.freq$number)

ggplot(data = obese_age.freq, aes(x = obese_age, y = proportion, fill=obese_age)) + geom_col()+
  labs(x="Age Group",y="Proportion of Obesity",title="Proportion of Obesity by Age Group")

```

@fig-scat1 displayed the proportion of obesity over each age group. 45-54 group occupied the highest proportion which was nearly 20% . 16-24 and 75+ groups shared the similar proportion which was approximately 8%. We can also see that the proportions per age appear to be normally distributed centred around 45-54 years of age suggesting those who are middle aged are more likely to be obese compared to their younger and older counter parts.

```{r}
#| echo: false
#| warning: false
#| fig-cap: Proportion of Obesity by Sex.
#| label: fig-scat2
#| fig-align: center
#| fig-width: 4.5
#| fig-height: 3.5
#gender
obese_sex.freq=table(data$Sex)
obese_sex.freq=as.data.frame(obese_sex.freq)
colnames(obese_sex.freq)=c("Sex","number")
obese_sex.freq$proportion=obese_sex.freq$number/sum(obese_sex.freq$number)

ggplot(data = obese_sex.freq, aes(x = Sex, y = proportion, fill=Sex)) + geom_col()+
  labs(x="Sex",y="Proportion of Obesity",title="Proportion of Obesity by Sex")
```

@fig-scat2 displays the proportion of obesity by sex. Females occupied a higher proportion of obesity which was nearly 55% . Whereas males occupied a lower proportion of obesity which was nearly 45%.

```{r}
#| echo: false
#| warning: false
#| fig-cap: Proportion of Obesity by Employment status.
#| label: fig-scat3
#| fig-align: center
#| fig-width: 4.5
#| fig-height: 3.5

#socio-economic status
obese_employment.freq <- table(data$Employment)
obese_employment.freq <- as.data.frame(obese_employment.freq)
obese_employment.freq$Var1 <- as.character(obese_employment.freq$Var1)
obese_employment.freq[3, 1] <- "Employed or on gov't training"

colnames(obese_employment.freq) <- c("obese_employment","number")
obese_employment.freq$proportion <- obese_employment.freq$number/sum(obese_employment.freq$number)


ggplot(data = obese_employment.freq, aes(x = obese_employment, y = proportion, fill=obese_employment)) + 
  geom_col()+
  labs(x="Employment Status",y="Proportion of Obesity",title="Proportion of Obesity by Employment Status")+
  coord_flip()+
    theme(legend.position = "none")

```

@fig-scat3 displays the proportion of obesity over each socio-economic status. The most obese socio-economic group was "In paid employment, self-employed or on gov't training" which was occupied by the highest proportion of obese observations, which was nearly 55.55% . Whereas "Doing something else" had the lowest proportion of obese observations which was approximately 1%.


# Formal Analysis {#sec-FA}
=======
# Formal Analyis {#sec-FA}

After conducting some exploratory analysis to help us visualise the data in @sec-EA, we now conduct some more rigorous analysis by fitting a series of logistic regression models.

We first fit a model that took into account all predictor variables and interaction terms. After following a process of backwards elimination we determined that the interaction terms were not significant and by following the Principle of Occam's Razor, we proceeded to fit the most simple model.


## Obese by Sex

=======

## Obesity by Sex

```{r}

#| echo: false
#| label: tbl-sex
#| tbl-cap: Estimates of the regression model coefficients.

modelsex <- glm(Obese ~ Sex,
                data = data, family = binomial)

```

| term      | estimate | p-value        |
|-----------|----------|----------------|
| intercept | -0.85652 | \<2e-16 \*\*\* |
| SexMale   | -0.02106 | 0.448          |

: Regression output for obesity by sex {#tbl-sexreg}

From the regression output in @tbl-sexreg we get the following regression equation

$$ \mbox{ln}({\frac{p}{1-p}}) = \alpha + \beta \cdot\mbox{Sex}= -0.86 -0.02 \cdot \mathbb{I}_{\mbox{Sex}}(\mbox{Male})$$

The baseline category for the binary response is not obese and the baseline category for sex is male. Therefore the estimates from the logistic regression model above represent a change on the log-odds scale for obesity compared to the response baseline of not obese for females in comparison to males.

From our regression output we can also determine the probability that someone is obese given that they are male and then the probability that someone is obese given that they are female which is determined to be the following:

```{r}

mod.sex.coef.logodds <- modelsex |>
                            summary() |>
                            coef()
probm <- plogis(mod.sex.coef.logodds["(Intercept)", "Estimate"])

```

```{r}
#| echo: false
probf <- plogis(mod.sex.coef.logodds["(Intercept)", "Estimate"] +
         mod.sex.coef.logodds["SexMale", "Estimate"])
```

| Sex    | Probability |
|--------|-------------|
| Male   | 0.2980659   |
| Female | 0.293679    |

: Probability of obesity by sex {#tbl-probsex}

```{r}
#| echo: false
#| label: fig-sex
#| tbl-cap: Confidence Interval Plot for Sex
data <- data |>
                  mutate(probs.sex = fitted(modelsex))
plot_model(modelsex, type = "pred", terms = "Sex", axis.title = c("Gender", "Prob. of Observation being Obese"), title = " ")
```

From these two values in @tbl-probsex we see that there is not much of a difference between probability that someone is obese given they are male compared to the probability that someone is obese given they are female. This coincides with the regression output in @tbl-sexreg as the p-value for males is not statistically significant. We are inclined to believe that gender is not an influential factor on obesity.

We also see that the difference in the null and residual deviance for the sex model is 1 and and this is smaller than the 95th percentile of a chi-squared distribution on 1 degree of freedom which is 3.84. Therefore it is not worth including sex in the model.

## Obesity by Age

```{r}

#| echo: false

modelage <- glm(Obese ~ AgeGroup,
                data = data, family = binomial)

```

=======
| Estimate | p-value  |
|----------|----------|
| -1.73631 | \<2e-16  |
| 0.44453  | 2.68e-09 |
| 0.88317  | \<2e-16  |
| 0.95485  | \<2e-16  |
| 1.16308  | \<2e-16  |
| 1.11899  | \<2e-16  |
| 0.81453  | \<2e-16  |

: Estimate of Regression Coefficients and P-values


The difference in the null and residual deviance for the age model is 499 and this is much larger than the 95th percentile of a chi-squared distribution on 6 degrees of freedom which is 12.59. Therefore it is worth including age in the model.

The baseline category for the binary response is not obese and the baseline category for age is 16-24. This means that estimates from the logistic regression model are for a change on the log-odds scale for being obese in comparison to the response baseline of not obese and for all other age groups in comparison to age group 16-24.

The model equation is

$$
\mbox{ln}({\frac{p}{1-p}}) = \alpha + \beta \cdot\mbox{Age Group}= -1.73 + 0.44 \cdot \mathbb{I}_{\mbox{Age Group}}(\mbox{25-34}) + 0.88 \cdot \mathbb{I}_{\mbox{Age Group}}(\mbox{35-44}) + 0.95 \cdot \mathbb{I}_{\mbox{Age Group}}(\mbox{45-54}) + 1.16 \cdot \mathbb{I}_{\mbox{Age Group}}(\mbox{55-64})+ 1.11 \cdot \mathbb{I}_{\mbox{Age Group}}(\mbox{65-74}) + 0.81 \cdot \mathbb{I}_{\mbox{Age Group}}(\mbox{75+})
$$

where $\mathbb{I}_{\mbox{Age Group}}$ is an indicator function For example the log odds increase of being obese increase by 0.44 for people in the 25-34 age group.

The probabilities of being obese for each age group starting at age group 16-24 are

```{r}
#| echo: false
modelage.coef.logodds <- modelage |>
                            summary() |>
                            coef()
```

```{r}
#| echo: false
#| eval: false
plogis(modelage.coef.logodds["(Intercept)", "Estimate"])

plogis(modelage.coef.logodds["(Intercept)", "Estimate"] +
         modelage.coef.logodds["AgeGroup25-34", "Estimate"])
plogis(modelage.coef.logodds["(Intercept)", "Estimate"] +
         modelage.coef.logodds["AgeGroup35-44", "Estimate"])
plogis(modelage.coef.logodds["(Intercept)", "Estimate"] +
         modelage.coef.logodds["AgeGroup45-54", "Estimate"])
plogis(modelage.coef.logodds["(Intercept)", "Estimate"] +
         modelage.coef.logodds["AgeGroup55-64", "Estimate"])
plogis(modelage.coef.logodds["(Intercept)", "Estimate"] +
         modelage.coef.logodds["AgeGroup65-74", "Estimate"])
plogis(modelage.coef.logodds["(Intercept)", "Estimate"] +
         modelage.coef.logodds["AgeGroup75+", "Estimate"])
```

```{r}

```

| Age Group | Probability |
|-----------|-------------|
| 16-24     | 0.1498      |
| 25-34     | 0.2156      |
| 35-44     | 0.2989      |
| 45-54     | 0.3140      |
| 55-64     | 0.3605      |
| 64-75     | 0.3504      |
| 75+       | 0.2846      |

: Probability of Obesity by Age Group

We can see the probability of being obese is highest for the 55-64 age group and lowest for the 16-24. We can plot the estimated probabilities by age group.

```{r}

plot_model(modelage, type = "pred", terms = "AgeGroup", axis.title = c("Age Group", "Probability of Obesity"), title = "Probability of Obesity by Age Group ")
```

We can see that there is quite a lot of difference between the estimated probabilities of being obese in each group.


## Obesity by Employemnt Status


```{r}
modelecon <- glm(Obese ~ Employment,
                 data = data, family = binomial)

```

The difference in the null and residual deviance for the socio-economic status model is 277 and and this is larger than the 95th percentile of a chi-squared distribution on 6 degrees of freedom which is 12.59. Therefore it is worth including socio-economic status in the model.

For socio-economic status we have the following model,

$$
ln(\frac {p} {1-p}) = \alpha + -1.16 \cdot \mathbb{I}_{\mbox{Edu}}(x) + -0.03 \cdot \mathbb{I}_{\mbox{Emp}}(x) + -0.06 \cdot 
\mathbb{I}_{\mbox{Home}}(x) + -0.12 \cdot \mathbb{I}_{\mbox{LFW}}(x) + 0.42 \cdot \mathbb{I}_{\mbox{Unable}}(x) + 0.17 \cdot \mathbb{I}_{\mbox{Retired}}(x)
$$

where $\mathbb{I}_{\mbox{Edu}}(x)$, $\mathbb{I}_{\mbox{Emp}}(x)$, $\mathbb{I}_{\mbox{Home}}(x)$, $\mathbb{I}_{\mbox{LFW}}(x)$, $\mathbb{I}_{\mbox{Unable}}(x)$, $\mathbb{I}_{\mbox{Retired}}(x)$ are indicator functions such that: 'Edu' represents subjects in full-time education, 'Emp' represents paid employment, self-employment or government training, 'Home' represents looking after home or family, 'LFW' represents looking for work, 'Unable' represents unable to work and 'Retired' represents subjects that are retired.

We again have that the baseline category for the binary response is 'not obese' and the baseline for employment status is 'Doing Something Else'. This means we can interpret the estimates from the logistic regression as the change in the log-odds scale for 'obese' compared to the baseline 'not obese' and for all other employment status in comparison to 'Doing Something Else'.


The probabilities of being obese for each socio-economic status are


```{r}
modelecon.coef.logodds <- modelecon |>
                            summary() |>
                            coef()
```

```{r}
plogis(modelecon.coef.logodds["(Intercept)", "Estimate"])

plogis(modelecon.coef.logodds["(Intercept)", "Estimate"] +
         modelecon.coef.logodds["EmploymentIn full-time education", "Estimate"])
plogis(modelecon.coef.logodds["(Intercept)", "Estimate"] +
         modelecon.coef.logodds["EmploymentIn paid employment, self-employed or on gov't training", "Estimate"])
plogis(modelecon.coef.logodds["(Intercept)", "Estimate"] +
         modelecon.coef.logodds["EmploymentLooking after home/family", "Estimate"])
plogis(modelecon.coef.logodds["(Intercept)", "Estimate"] +
         modelecon.coef.logodds["EmploymentLooking for/intending to look for paid work", "Estimate"])
plogis(modelecon.coef.logodds["(Intercept)", "Estimate"] +
         modelecon.coef.logodds["EmploymentPerm unable to work", "Estimate"])
plogis(modelecon.coef.logodds["(Intercept)", "Estimate"] +
         modelecon.coef.logodds["EmploymentRetired", "Estimate"])
```

Here we see that the probability of being obese is highest for subjects that are permanently unable to work, it is lowest for subjects in full-time education.

```{r}
plot_model(modelecon, type = "pred", terms = "Employment", axis.title = c("Socio-economic Status", "Probability of Obesity"), title = "Probability of Obesity by Socio-economic Status")
```

From the plot we can see there appears to be a significant difference in the probability of being obese between different socio-economic groups.
