---
title: "Group_8_Formal_Analysis"
echo: false
format: html
editor: visual
self-contained: true
---

```{r}


data <- read.csv("DAProject9.csv")
library(dplyr)
data <- data %>%
  mutate(Obese = ifelse(BMIgroup == "Obese", 1, 0))
library(ggplot2)
library(tidyverse)
library(tidymodels)
library(gt)
library(MASS)
library(patchwork)
library(moderndive)
```

```{r}
#fitting a logistic regression model for each explanatory variable 


#age
modelage <- glm(Obese ~ AgeGroup,
                data = data, family = binomial)
summary(modelage)

#sex
modelsex <- glm(Obese ~ Sex,
                data = data, family = binomial)
summary(modelsex)

#socio-economic status
modelecon <- glm(Obese ~ Employment,
                 data = data, family = binomial)
summary(modelecon)

```

```{r}

```

The baseline category for the binary response is not obese and the baseline category for age is 16-24. This means that estimates from the logistic regression model are for a change on the log-odds scale for being obese in comparison to the response baseline of not obese and for all other age groups in comparison to age group 16-24.

$$ \mbox{ln}({\frac{p}{1-p}}) = \alpha + \beta \cdot\mbox{Sex}= -0.86 -0.02 \cdot \mathbb{I}_{\mbox{Sex}}(\mbox{Male})$$

As before, the baseline category for the binary response is not obese and the baseline category for sex is male. Therefore the estimates from the logistic regression model above represent a change on the log-odds scale for obesity compared to the response baseline of not obese for females in comparison to males.

From our regression output we can also determine the probability that someone is obese given that they are male which is determined to be the following:

```{r}
mod.sex.coef.logodds <- modelsex |>
                            summary() |>
                            coef()
plogis(mod.sex.coef.logodds["(Intercept)", "Estimate"])
```
And the probability that someone is obese given they are female. 
```{r}
plogis(mod.sex.coef.logodds["(Intercept)", "Estimate"] +
         mod.sex.coef.logodds["SexMale", "Estimate"])
```

From these two values we see that there is not much of a difference between probability that someone is obese given they are male compared to the probability that someone is obese given they are female. This coincides with the regression output about @tbl-sex as the p-value for males is not statistically significant. We are inclined to believe that gender is not an influential factor on obesity.
